---
title: "Hands-on Ex 2: Beyond ggplot2 Fundamentals"
author: "Liang Xiuhao Rydia"
date: "April 20, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  warning: false
  freeze: true
---

## 1 Overview

This chapter / Hands-on exercise is an introduction to several ggplot2 extension. Key idea is to create more elegant and effective statistical graphics, pleasing to the eyes of the reader, and reducing the cognitive load of the reader.

Key learning:

-   control the placement of annotation on a graph by using functions provided in ggrepel package

-   create professional publication quality figure by using functions provided in ggthemes and hrbrthemes packages

-   plot composite figure by combining ggplot2 graphs by using patchwork package.

## 2 Getting started

::: panel-tabset
## Installing and loading the required libraries

In this exercise, beside tidyverse, four R packages will be used. They are:

-   ggrepel: an R package provides geoms for ggplot2 to repel overlapping text labels.
-   ggthemes: an R package provides some extra themes, geoms, and scales for ‘ggplot2’.
-   hrbrthemes: an R package provides typography-centric themes and theme components for ggplot2.
-   patchwork: an R package for preparing composite figure created using ggplot2. 

Code chunk below will be used to check if these packages have been installed and also will load them onto your working R environment.

```{r}
pacman::p_load(ggrepel,patchwork,
               ggthemes,hrbrthemes,
               tidyverse)
```

## Importing data

For the purpose of this exercise, a data file called Exam_data will be used. It consists of year end examination grades of a cohort of primary 3 students from a local school. It is in csv file format.

The code chunk below imports exam_data.csv into R environment by using [read_csv()](https://readr.tidyverse.org/reference/read_delim.html) function of [**readr**](https://readr.tidyverse.org/) package. **readr** is one of the tidyverse package.

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## glimpse() the Data

```{r}
glimpse(exam_data)
```

There are a total of seven attributes in the exam_data tibble data frame. Four of them are categorical data type and the other three are in continuous data type.

-   Categorical attributes: ID, CLASS, GENDER and RACE.

-   Continuous attributes: MATHS, ENGLISH and SCIENCE.
:::

## 3 Beyond ggplot2 Annotation: ggrepel

Annotation for large number of data points is potentially challenging, as the annotations may overwhelm the plot. In this example, `geom_label()` is used to show the labels on all the data points. The resulting plot below shows the annotation covering much of the data points. 

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              linewidth=0.5) +  
  geom_label(aes(label = ID), 
             hjust = 0.5, 
             vjust = 0.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## The code

```{r}
#| eval: false
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              linewidth=0.5) +  
  geom_label(aes(label = ID), 
             hjust = 0.5, 
             vjust = 0.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::

[ggrepel](https://ggrepel.slowkow.com/) is an extension of **ggplot2** package which provides `geoms` for **ggplot2** to repel overlapping text as in our example above.

We simply replace `geom_text()` by [`geom_text_repel()`](https://ggrepel.slowkow.com/reference/geom_text_repel.html) and `geom_label()` by [`geom_label_repel`](https://ggrepel.slowkow.com/reference/geom_text_repel.html).

::: panel-tabset
## The plot
```{r}
#| echo: False
ggplot(exam_data,
       aes(x= MATHS,
       y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm,
              linewidth = 0.5) +
  geom_label_repel(aes(label = ID),
                   hjust = 0.5,
                   vjust = 0.5) +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

```

## The code
```{r}
#| eval: False
ggplot(exam_data,
       aes(x= MATHS,
       y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm,
              linewidth = 0.5) +
  geom_label_repel(aes(label = ID),
                   hjust = 0.5,
                   vjust = 0.5) +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

```
:::

## 4 Beyond ggplot2 themes

