---
title: "Take-home Ex 01"
author: "Liang Xiuhao Rydia"
date: "April 15, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  warning: false
  freeze: true
---

# **Creating data visualisation beyond default**

## 1.0 Setting the scene

There are two major residential property market in Singapore, namely public and private housing. Public housing aims to meet the basic need of the general public with monthly \$14,000 for families/couples buying a new HDB flat. Singapore is a city-state with limited land resource. Housing, on the other hand, is one of the basic need of people of this city-state. In order to ensure that all citizen in the country can access to affordable housing, the government have initiated massive public housing programme since independent.

## 2.0 The Task

Assuming the role of a graphical editor of a median company, you are requested to prepare minimum two and maximum three data visualisation to reveal the ***private residential market*** and ***sub-markets of Singapore*** for the 1st quarter of 2024.

```{r}
# load tidyverse from pacman
pacman::p_load(tidyverse,dplyr)

# load transaction data using readr, part of tidyverse package
data <- read_csv("data/ResidentialTransaction20240414220633.csv")

```

```{r}
# have a basic understanding for all 21 variables
glimpse(data)

```

```{r}
# have a basic understanding for all 21 variables
glimpse(data)

```



```{r}
ggplot(data = data, 
       aes(x=`Type of Sale`)) +
  geom_bar() 

```
```{r}
ggplot(data = data, 
       aes(x=`Type of Area`)) +
  geom_bar() 

```

```{r}
ggplot(data = data,
       aes(x =`Planning Area`,
           fill = `Type of Sale`)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

