---
title: "Take-home Ex 01"
author: "Liang Xiuhao Rydia"
date: "April 15, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  warning: false
  freeze: true
---

# **Creating data visualisation beyond default**

## 1.0 Overview

### 1.1 Setting the scene

There are two major residential property market in Singapore, namely public and private housing. Public housing aims to meet the basic need of the general public with monthly \$14,000 for families/couples buying a new HDB flat. Singapore is a city-state with limited land resource. Housing, on the other hand, is one of the basic need of people of this city-state. In order to ensure that all citizen in the country can access to affordable housing, the government have initiated massive public housing programme since independent.

### 1.2 The Task

Assuming the role of a graphical editor of a media company, you are requested to prepare minimum two and maximum three data visualisation to reveal the ***private residential market*** and ***sub-markets of Singapore*** for the 1st quarter of 2024.

## 2.0 Set Up

### 2.1 Loading the packages into R

T he `pac::p_load()` function to load the required R packages in the working environment. The following packages are used in this set up:

-   ggthemes: Extra themes, geoms, and scales for ggplot2.
-   tidyverse: A collection of core packages designed for data science, used extensively for data preparation and wrangling.
-   To add XXX

```{r}
# load tidyverse from pacman
pacman::p_load(tidyverse,dplyr)
```

### 2.2 Importing the data sets

Five sets of data are provided for this exercise:

1.  2023 Quarter 1 Residential Transaction: ResidentialTransaction20240308160536.csv
2.  2023 Quarter 2 Residential Transaction: ResidentialTransaction20240308160736.csv
3.  2023 Quarter 3 Residential Transaction: ResidentialTransaction20240308161009.csv
4.  2023 Quarter 4 Residential Transaction: ResidentialTransaction20240308161109.csv
5.  2024 Quarter 1 Residential Transaction: ResidentialTransaction20240414220633.csv

All five sets of data consists of past transactions between Q1 of 2023 to Q1 of 2024, with 21 variables that will be explored in detail after the data sets are imported.

```{r}
# load transaction data using readr, part of tidyverse package
data.23Q1 <- read_csv("data/ResidentialTransaction20240308160536.csv")
data.23Q2 <- read_csv("data/ResidentialTransaction20240308160736.csv")
data.23Q3 <- read_csv("data/ResidentialTransaction20240308161009.csv")
data.23Q4 <- read_csv("data/ResidentialTransaction20240308161109.csv")
data.24Q1 <- read_csv("data/ResidentialTransaction20240414220633.csv")
```

To have a basic understanding of all 21 variables, `glimpse()` is used with 2024 Q1 data. We can see that there are 21 coloumns, with the following data types:

-   doubles:

    1.  `Transacted Price ($)`
    2.  `Area (SQFT)`
    3.  `Unit Price ($ PSF)`
    4.  `Area (SQM)`
    5.  `Unit Price ($ PSM)`
    6.  `Number of Units`

-   characters: the rest of the 15 variables

```{r}
# have a basic understanding for all 21 variables
glimpse(data.24Q1)

```



```{r}
ggplot(data = data.24Q1, 
       aes(x=`Type of Sale`)) +
  geom_bar() 

```

```{r}
ggplot(data = data.24Q1, 
       aes(x=`Type of Area`)) +
  geom_bar() 

```

```{r}
ggplot(data = data.24Q1,
       aes(x =`Planning Area`,
           fill = `Type of Sale`)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```
